### 2014-05-01 paper.tex
Added final changes

\textbf{A Major Qualifying Project\\
Submitted to the Faculty of\\
Worcester Polytechnic Institute\\
in partial fulfillment of the requirements for the\\
Degree in Bachelor of Science\\
in\\
Mathematical Sciences
and
Computer Science\\
By
\\*[12pt]}

@@

In this paper we discuss Worcester Polytechnic Institute and a team at Carnegie Mellon University use the Boston Dynamics Atlas robot at the DARPA Robotics Challenge (DRC). Initially we will discuss about the tasks involved in the challenge, and the hardware specifications and limitations. We will then progress to the software architecture and the mathematics behind the controller.

@@

@@

@@

Any efficient computational approach must work with an approximate state space and an approximate controller and an approximation to the solution to the dynamic programming equations.  Atkison's \cite{random_sampling_dynamic_programming} and his co-authors have studied an approximation method that replaces a full optimization at each point in the state space with a simple comparison between the current “best” control strategy and a single, ramdomly generated, alternative.  Whitman \cite{eric_thesis} used this approximation technique but also studies an approach that separates the high-dimensional control problem for the Atlas robot into a family of low-dimensional (and so feasible) optimization problems and builds an approximate solution by integrating information from the many separate problems.  
Part of the work done in this project was on the implementation of approximation methods and techniques for speeding up the dynamic programming algorithm.  The approach is applied to the simplest model problem: the swing-up pendulum. \cite{eric_thesis}

@@

For external sensors Atlas has:

@@

The MultiSense SL is a tri-modal range sensor, priving laser range data, stereo range data, and video output stream. The device provides the majority of perceptual data used for teleoperation as well as automated control.

@@

Two Situational awareness cameras are placed just below the head of the robot, providing a wide range view of the workspace.

@@

The robot gives the option of interchanging the hands. DARPA offered two primary options: the iRobot© hands, and the Sandia© hands. While both options are from known companies in the industry, neither provided the power, nor robustness necessary to complete the high demanding tasks of the DRC. These hands were designed to perform precision tasks, such as picking up a coin, or a fist-sized rock, but not to handle a power tool, or hold the robot while climbing up a ladder. The team chose an alternative end-effector developed by Robotiq©, a 3-Finger adaptive industrial robot gripper, which provided a much stronger grip.

@@

\subsection{DBI Behaviors}
Boston Dynamics provided predefined behaviours for Atlas that allow for basic performance like balancing and walking. Transitions between behaviors can occur manually or automatically, and follow a pre-determined state-machine.

\subsubsection{Manipulate}
\begin{wrapfigure}{r}{0.3\textwidth}
  \begin{center}
    \includegraphics[scale=0.5]{images/align_to_wall.png}
  \end{center}
  \caption{Rviz within ROS}
\end{wrapfigure}
The Manipulation mode, in which the user can control all the joints above the torso, and basic control of the pelvis orientation, and the pelvis height. The legs are controlled by the Boston Dynamics controller embedded in the firmware.

The manipulation mode has a wide range of motion if the pelvis position and orientation is controlled together with an inverse kinematics solver for the joints, but can be unpredictable. The robot has the problem that the back joint that helps torso move back and forth is not strong enough to hold, and may give out a times, causing the whole top part of the robot fall forward, or at times, make the entire robot fall forward.

The user has the option to input any additional weight that the hand is carrying to help the controller.

@@

Boston Dynamics walking behaviour works by the user providing a sequence of steps paramters, which include the step duration, sway duration, swing height, lift height, and step end distance.

The WPI team developed an additional interface for Rviz, such as the robot driver is able to specify a final desired feet positions for the robot, and the software produces all the intermediate steps. It took some trial-and-error testing to figure out what the maximum step size are possible, and the program does not account for any debris on the floor.

@@

The system was setup in three major components:

@@

\subsubsection{Field Computer}
The Field Computer is designed to handle the team programmed controller logic of the robot, either being a full body controller, or using the predefined behaviors given by Boston Dynamics.

@@

Also responsible for sending and receiving high level commands from the Operational Command Unit.

@@

There can be more than one OCU on a team, and it is the interface to which the operator has to evaluate the robot situation, and perform the desired tasks.

@@

DARPA structured a predictable, but limited bandwidth communication between the field computer and the OCU. The network quality varied every 60 seconds between a $1000 \frac{kb}{s}$ with a $100ms$ round trip delay, and a $100 \frac{kb}{s}$ with a $500ms$ round trip delay. This setup is to incentivize the teams to make the robot perform more autonomous tasks.

@@

\subsection{SD/FAST}
The library SD/Fast provides physically-based simulation of mechanical systems by taking a short description of an articulated system of rigid bodies and deriving the full nonlinear equations of the system. \cite{sdfast} The library takes several key factors about each sub-body in the robot, such as the link positions, mass, and inertia. The library than generates a C file that provides information such as the center of mass, jacobians, and positions.

@@

The robot is defined as tree/skeleton starting from the pelvis, defining the weight and distances between each joint. When we generate Jacobians \cite{robot_modeling} to move the end effector of each limbs, we also give it the ability to move the pelvis on a global coordinate frame. The hand jacobians has no direct effect on the the leg joints, but indirectly moves the imaginary point that the pelvis should follow, and the legs move with it.

@@

Usually there are about 36 active jacobians, there may be more or less depending on the required constraints, such as disabling a hand constraint, or enabling a desired elbow position constraint. The next procedure is an algorithm that is able to intake all the competing needs, and output the desired joint velocities.

The controller uses a system of calculating a pseudo-inverse, and finding a least squares solution by using quadratic programming. We first generate a matrix, $A$, of 34 columns and about 68 rows, and a vector, $b$, with the same name of rows. The matrix is composed of all the jacobians multiplies by an weight, that must be equal to the desired change. An additional regularization identity element is added to help ensure that none of the joints move too quickly.

@@

Several unexpected behaviors showed from this experiment: The controller would prefer to to leave joints at the joint limits unless it was complete necessary to move it to achieve a necessary motion, and when moving the arm away from the ground, the controller will choose to move the legs first until the legs reach a singularity.

@@

\subsection{Results}
Overall the controller is able to perform full body motions without much work from the developer, and sometimes make almost human-like behavior. Where used, it performed well on the DRC trials, and gave little trouble for the operator.

@@

The initial approach to complete this task was to use the Boston Dynamics motion controller, and perform the motion by solving for different positions using MoveIt!, but this approach quickly presented several problems, such as when solving for the 6-dof arm, it would only occasionally find a solution, and it was unable to complete the full motion. Another problem arised when using the 6-dof arm + 3-dof back joints, MoveIt! would use the back joints way too much for the motions and cause unpredictable behaviour.

@@

The left hand was accidentally turned on into relative positions, and since there were no graphical feedback about the operation of each hand, the operator was unable to determine why the robot was unable to move as freely as necessary to cut the wall. On an attempt to reset the controller while the robot was still running to fix the unexpected behavior, the robot fell, having to reset the task.

@@

The expected behavior for the control law is to apply a torque equal to the direction the pendulum is swinging, such that more energy is added to the pendulum with each period, and when the pendulum finally has enough energy to perform a full rotation, apply a torque opposite to the movement such that the pendulum stops in an upright position.

@@

We run the previous operation on all states of the grid for an uncertain number of times. For this particular program, running this optimization for 1000 iterations is siffucient for making the pendulum converge to the desired goal state, but it can also be noted that running for 10000 times generates a more torque efficient solution.

@@

\section{Analysing this solution}
The algorithm is far from being real time, since for the single-link pendulum case, on an 8-core CPU at 2.8GHz, takes about 4 seconds for each iteration, and about 1 hour for the 1000 iterations, but it does generate very robust policy tables.

On the following plot we show the policy of 10,000 iterations running against a pendulum, position of the pendulum relative to the up position, starting at pi, pointing down, and with 0 velocity. It is interesting to see that the policy follows similar to a bang-bang controller, using near maximum torque on both directions, until the system has enough energy to reach the necessary solution. (TODO: Talk about behavior near t=45seconds, where the exact amount of force is applied to get the up position)

@@

@@

There the team has performed well on the DRC trials, it is important as well to create a strong code base that can evolve over the years, and adapt to future challenges. An important aspect of having a strong base is being able to perform experiments, and build upon previously developed work.

>>>>

\textbf{A Major Qualifying Project\\
Submitted to the Faculty of\\
Worcester Polytechnic Institute\\
in partial fulfilment of the requirements for the\\
Degree in Bachelor of Science\\
in\\
Mathematical Sciences
and
Computer Science\\
By
\\*[12pt]}

@@

This paper presents the Worcester Polytechnic Institute(WPI) and Carnegie Mellon University(CMU) team approach for competing on the DARPA Robotics Challenge (DRC) using the humanoid Boston Dynamics Atlas robot. An overview and analysis of the hardware and software architecture is described with emphasis on two of the challenges tasks, Wall and Drill. The realization of a double stance Inverse Kinematics(IK) full-body controller used for manipulation and its overall performance is elaborated. Moreover, an analysis for using Dynamic Programming optimization for robotic humanoid path planning is developed.

@@

Special thanks for Worcester Polytechnic Institutes, and the great support it has had on my education, and the influence on my career. Honourable mentions towards Mathew DeDonato, Velin Dimitrov, and Perry Franklin.

A huge thanks for my brother, Felipe Polido, to have introduced me into the project, and helped me with the many challenges of participating in a team, working through problems, and finding new solutions.

@@

\begin{center}
\includegraphics[scale=0.1]{images/atlas3.jpg}
\end{center}

@@

Any efficient computational approach must work with an approximate state space and an approximate controller and an approximation to the solution to the dynamic programming equations.  Atkison's \cite{random_sampling_dynamic_programming} and his co-authors have studied an approximation method that replaces a full optimization at each point in the state space with a simple comparison between the current “best” control strategy and a single, randomly generated, alternative.  Whitman \cite{eric_thesis} used this approximation technique but also studies an approach that separates the high-dimensional control problem for the Atlas robot into a family of low-dimensional (and so feasible) optimization problems and builds an approximate solution by integrating information from the many separate problems.  
Part of the work done in this project was on the implementation of approximation methods and techniques for speeding up the dynamic programming algorithm.  The approach is applied to the simplest model problem: the swing-up pendulum. \cite{eric_thesis}

@@

Atlas is equipped with the following perception sensors:

@@

The MultiSense SL is a tri-modal range sensor, providing laser range data, stereo range data, and video output stream. The device supply the majority of perceptual data used for tele-operation as well as automated control.

@@

Two situational awareness cameras are placed just below the head of the robot, providing a wide range view of the workspace.

@@

The robot gives the option of interchanging the hands. DARPA offered two primary options: the iRobot© hands, and the Sandia© hands. While both options are from known companies in the industry, neither provided the power, nor robustness necessary to complete the high demanding tasks of the DRC. These hands were designed to perform precision tasks, such as picking up a coin, or a fist-sized rock, but not to handle a power tool, or hold the Atlas robot while climbing up a ladder. The team chose an alternative end-effector developed by Robotiq©, a 3-Finger adaptive industrial robot gripper, which provided a much stronger grip.

@@

\subsection{DBI Behaviours}
Boston Dynamics provided predefined behaviours for Atlas that allow for basic performance like balancing and walking. Transitions between behaviours can occur manually or automatically, and follow a pre-determined state-machine.

\subsubsection{Manipulate}
\begin{wrapfigure}{r}{0.3\textwidth}
  \begin{center}
    \includegraphics[scale=0.5]{images/align_to_wall.png}
  \end{center}
  \caption{Rviz within ROS}
\end{wrapfigure}
In manipulation mode the user can control all the joints above the torso, and basic control of the pelvis orientation and height. The legs and overall balance are controlled by the Boston Dynamics controller embedded in the firmware.

The manipulation mode has a wide range of motion if the pelvis position and orientation is controlled together with an inverse kinematics solver for the joints, but can be unpredictable. The robot back joint that controls forwards and backwards motion was originally to weak, is unable to maintain position in certain configurations. The failure of this joint may compromise any given manipulation task.

Lastly, in this mode the user has the option to input any additional weight from objects in the hand to help the controller maintain balance.

@@

Boston Dynamics walking behaviour works by the user providing a sequence of steps parameters, which include the step duration, sway duration, swing height, lift height, and step end distance.

The WPI team developed an additional interface for Rviz so that the robot operator is able to specify a final desired feet positions for the robot, and the software produces all the intermediate steps. It took some trial-and-error testing to figure out what the maximum step size is, and the program does not account for any debris on the floor.

@@

The system architecture was setup in three major components:

@@

\subsubsection{Field Computer}
The Field Computer is designed to handle the team programmed controller logic of the robot, either being a full body controller, or using the predefined behaviours given by Boston Dynamics. This is also where high-bandwidth perception and autonomous algorithms function.

@@

Lastly, it is responsible for sending and receiving high level commands from the Operational Command Unit.

@@

The Operational Command Unit(OCU) is the interface between the operator and the field computer. Our team relied on multiple operators working in individual OCU computers networked together. From the OCU the operators have to evaluate the robot situation, reason, and execute the desired tasks.

@@

DARPA structured a predictable, but limited bandwidth communication between the field computer and the OCU. The network quality varied every 60 seconds between a $1000 \frac{kb}{s}$ with a $100ms$ round trip delay, and a $100 \frac{kb}{s}$ with a $500ms$ round trip delay. This setup is to incentive the teams to perform the tasks autonomously.

@@

\subsection{SD/FAST}
The library SD/Fast provides physically-based simulation of mechanical systems by taking a short description of an articulated system of rigid bodies and deriving the full non-linear equations of the system. \cite{sdfast} The library takes several key factors about each sub-body in the robot, such as the link positions, mass, and inertia. The library than generates a C file that provides information such as the center of mass, Jacobian, and positions.

@@

The robot is defined as tree/skeleton starting from the pelvis, defining the weight and distances between each joint. When we generate Jacobians \cite{robot_modeling} to move the end effector of each limbs, we also give it the ability to move the pelvis on a global coordinate frame. The hand Jacobians has no direct effect on the the leg joints, but indirectly moves the imaginary point that the pelvis should follow, and the legs move with it.

@@

Usually there are about 36 active Jacobians, there may be more or less depending on the required constraints, such as disabling a hand constraint, or enabling a desired elbow position constraint. The next procedure is an algorithm that is able to intake all the competing needs, and output the desired joint velocities.

The controller uses a system of calculating a pseudo-inverse, and finding a least squares solution by using quadratic programming. We first generate a matrix, $A$, of 34 columns and about 68 rows, and a vector, $b$, with the same name of rows. The matrix is composed of all the Jacobians multiplies by an weight, that must be equal to the desired change. An additional regularization identity element is added to help ensure that none of the joints move too quickly.

@@

Several unexpected behaviours showed from this experiment: The controller would prefer to to leave joints at the joint limits unless it was complete necessary to move it to achieve a necessary motion, and when moving the arm away from the ground, the controller will choose to move the legs first until the legs reach a singularity.

@@

\subsection{Results}
Overall the controller is able to perform full body motions without much work from the developer, and sometimes act in an almost human-like behaviour. On the cases in which it was used, the full-body controller performed well on the DRC trials, and gave little trouble for the operator.

@@

The initial approach to complete this task was to use the Boston Dynamics motion controller, and perform the motion by solving for different positions using MoveIt!, but this approach quickly presented several problems, such as when solving for the 6-DoF arm, it would only occasionally find a solution, and it was unable to complete the full motion. Another problem arose when using the 6-DoF arm + 3-DoF back joints, MoveIt! would use the back joints way too much for the motions and cause unpredictable behaviour. There was no direct way of limiting the motion of specific joints through MoveIt!.

@@

The left hand was accidentally turned on into relative positions, and since there were no graphical feedback about the operation of each hand, the operator was unable to determine why the robot was unable to move as freely as necessary to cut the wall. On an attempt to reset the controller while the robot was still running to fix the unexpected behaviour, the robot fell, having to reset the task.

@@

The expected behaviour for the control law is to apply a torque equal to the direction the pendulum is swinging, such that more energy is added to the pendulum with each period, and when the pendulum finally has enough energy to perform a full rotation, apply a torque opposite to the movement such that the pendulum stops in an upright position.

@@

We run the previous operation on all states of the grid for an uncertain number of times. For this particular program, running this optimization for 1000 iterations is sufficient for making the pendulum converge to the desired goal state, but it can also be noted that running for 10000 times generates a more torque efficient solution.

@@

\section{Analysing this solution}
The algorithm is far from being real time, since for the single-link pendulum case, on an 8-core CPU at 2.8 GHz, takes about 4 seconds for each iteration, and about 1 hour for the 1000 iterations, but it does generate very robust policy tables.

On the following plot we show the policy of 10,000 iterations running against a pendulum, position of the pendulum relative to the up position, starting at pi, pointing down, and with 0 velocity. It is interesting to see that the policy follows similar to a bang-bang controller, using near maximum torque on both directions, until the system has enough energy to reach the necessary solution.

The following plot demonstrates the value and the policy values of the pendulum swing-up solution. The X-axis is the velocity of the pendulum, while the Y-axis is the current position of the pendulum.

Figure~\ref{fig:pendulum_plot} states that as the pendulum starts at $[\pi, 0]$, starting with the pointing down, with no velocity, and on the highest value on the value plot. Over time, the pendulum will travel down the apply torque and gain velocity over time, up to the point that the pendulum will reach the lowest value on the plot.

Figure~\ref{fig:pendulum_time_plot} shows the position, velocity, and torque being applied at any moment on the pendulum. Even with 1,000 iterations the system was able to find a method in which the position converges to 0, when the pendulum is pointing up, but one can also see that during the last 20 seconds, the line contains quite some noise.

Lastly, on Figure~\ref{fig:pendulum_plot2} shows the paths of pendulum. Each red line represents the start configuration of a pendulum, and stops when they go outside of the plot, or reaches the desired position at $[0,0]$. It was surprising to learn that on the top-right corner, when the pendulum has some momentum and is pointing down, they do not all converge exactly to the same path, but are rather somewhat parallel to one another.

\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{images/DDP.png}
\caption{Plot showing the value and policy of a pendulum swing up}
\label{fig:pendulum_plot}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=1.0]{images/Pendulum_time.png}
\caption{Position, velocity, and torque of pendulum over time}
\label{fig:pendulum_time_plot}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{images/direction_lines.png}
\caption{Paths that pendulum would follow given any start position}
\label{fig:pendulum_plot2}
\end{center}
\end{figure}

@@

\section{Results}
The results of the algorithm were never applied in a real system, and the generated value and policy for the GPU and CPU were essentially identical. Presenting sensor noise and other real-life factors into a simulation, we can still see that the system can still perform under inaccurate actuation. \cite{eric_thesis}

\section{Conclusion}
Dynamic Programming still has the Curse of Dimensionality, the dimensions of the project can be scaled down to fit into a manageable set, and since the data storage is rather comparably cheap, the information could be stored in disk for later retrieval. Furthermore, with the help of CUDA, specific sections of the grid could be recalculated with finer detail in real time when needed.

The Pendulum-Swing problem has been solved many times before, but with this type of solution, the problem can be extended to higher dimensions, and it only requires an linearised model of the system. The algorithm could be easily extended to take a double, triple, or any given number of links pendulum.

\section{Future Work}
This problem could be extended into looking at different end-effector positions for Atlas. One could generate a table of all feasible end-effector positions, and find paths on how to travel from one position to another while minimizing change in the center of mass deviation from the support polygon.

@@

The WPI-CMU team performed well on the DRC trials. Regardless, due to the time constraints a lot of the code was built in a hurry, and without a proper structure. It is important to create a strong code base that can evolve over the years, and adapt to future challenges. An important aspect of having a strong base is being able to perform experiments, and build upon previously developed work. A lot of emphasizes after the DRC trials was directed on refactoring the code base, and there is still a long way to go.
